Towards Automated Merging of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-05-06 sattose/paper.tex
Updates from Overleaf

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolpceurws}
\usepackage{float}
\usepackage{hyperref}
\usepackage{todonotes}

@@

\begin{abstract}
Duplication in source code can have a major negative impact on the maintainability of source code, as creates explicit dependencies between fragments of code. Typically between 10 and 20 percent of projects are cloned. In this study, we look into the opportunities to aid in the process of refactoring these duplication problems for object-oriented programming languages. To do this, we propose a method to be able to detect clones that are suitable for refactoring. This method focuses on taking into account the scope and coupling between expressions in code, to be sure our refactoring improves the design and does not leave side effects after it is applied.

@@

@@

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. For our tool, we decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio \cite{higo2013revisiting, kamalpriya2017enhancing}. However, instead of building a dependency graph, we build a similarity graph of statements linking to similar statements. We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is useable with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). This is displayed in figure \ref{fig:clonerefactor}.

@@

\subsubsection{Detection of clone classes}\label{chap:cloneclasses}
For CloneRefactor, we have chosen to focus on the detection of clone classes rather than clone pairs. We did this because clone pairs don’t provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. On the other hand, clone classes are harder to manage, but they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue involves clone grouping by pairs: single clone reference amount increases according to binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), of which a heavy information redundancy follows \cite{fontana2012duplicated}.

@@

\section{Redefining type 1, 2 and type 3 clones}
In a 2007 survey by Roy et al \cite{roy2007survey} he defined four types of clones:

@@

\subsection{Type 1 clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, does not yet indicate that they are actually identical. Even when textually equal, method calls can refer to different methods, type declarations can refer to different types and even variables can be of a different type. This could invalidate a refactoring opportunity for such a code fragment. Because of that, we recommend redefining type-1 clones for refactoring. We propose to:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} This way we can validate whether two method references, like method calls, are actually equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also of the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
\end{itemize}

>>>>

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{twocolpceurws}
\usepackage{float}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

@@

\begin{abstract}
Duplication in source code can have a major negative impact on the maintainability of source code, as it creates explicit dependencies between fragments of code. These explicit dependencies have been shown to often cause bugs, as a modification in one fragment might have consequences for the correctness of another. In this study, we look into the opportunities to aid in the process of refactoring these duplication problems for object-oriented programming languages. To do this, we propose a method to be able to detect clones that are suitable for refactoring. This method focuses on taking into account the scope and coupling between expressions in code, to be sure our refactoring improves the design and does not leave side effects after it is applied.

@@

\subsubsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be detected. In this study we focus on clone classes, because of the advantages for refactoring. Clone pairs don’t provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. Even though clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue involves clone grouping by pairs: single clone reference amount increases according to binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), of which a heavy information redundancy follows \cite{fontana2012duplicated}.

@@

\subsection{Survey on Clone Detection Tools}
We conducted a short survey on (recent) clone detection tools that we could use to analyze refactoring possibilities. We chose a set of tools that are open source and can analyze either Java or Python. Then we chose four criteria by which we considered these tools: 
\begin{enumerate}
    \item \textbf{Should find clones in any context.} We want to perform an analysis on a set of projects. In this analysis we want to be able to get a full overview of cloning in the set of projects. Some tools only find clones in specific contexts, like only method-level clones. Because of that, we want to use a clone detection tool that is able to find clones in any context.
\item \textbf{Should detect clone classes in a number of example projects.} We have created a number of test projects\footnote{Our test projects to assess the validity of clone detection tools: \url{https://github.com/SimonBaars/CloneRefactor/tree/master/src/test/resources}} that we used to assess the validity of clone detection tools. On basis of this, we checked whether clone detection tools can correctly find clones in diverse contexts.
\item \textbf{Considers dependency graphs to be sure that this clone can be merged.} When detecting clones for refactoring purposes, it is important that clone instances are actually equal. Sometimes, to verify actual equality (not just textual equality), it is required to consider resolved symbols (this is described more elaborately in section \ref{chap:type1clones}). Because of this, we want to use a clone detection tool that can analyze such structures.
\item \textbf{Makes it possible to configure the way in which clones are detected (as in what to exclude from the detection) in a relatively straightforward way.} We want to be able to exclude expressions/statements from matching (more about our rationale in section \ref{chap:clonetypes}). To be able to do this, the tool needs to be able to recognize the meaning of certain tokens.
\end{enumerate}

The results of our survey is displayed in table \ref{table:dettools}.

\begin{table}[H]
 \begin{center}
  \caption{Our survey on clone detection tools.} \label{table:dettools}
  \bigskip
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{Clone Detection Tool}                                                                                                                           & \textbf{Short Description}                                                          & \textbf{(1)}                  & \textbf{(2)} & \textbf{(3)} & \textbf{(4)} \\ \hline
Siamese \cite{ragkhitwetsagul2019siamese}                                                                                                              & \begin{tabular}[c]{@{}l@{}}ANTRL based\\ large scale clone\\ detection\end{tabular} &  &             &             & \checkmark            \\ \hline
NiCAD \cite{roy2008nicad, cordy2011nicad} \footnote{This clone detection tool has also been used by Fontana et al \cite{fontana2012duplicated}}      & \begin{tabular}[c]{@{}l@{}}Duplicated code\\ refactoring advisor\end{tabular}       & ?                             & ?            & ✗            & ✓            \\ \hline
CPD \cite{roy2009comparison}\footnote{CPD is a Clone Detection tool that is part of PMD. PMD is a widely used tool for static code analysis in Java.} & \begin{tabular}[c]{@{}l@{}}Widely used clone\\ detection tool\end{tabular}          & ✓                             & ✓            & ✗            & ✗            \\ \hline
\begin{tabular}[c]{@{}l@{}}CCFinder \cite{kamiya2002ccfinder}\\ CCFinderX\\ D-CCFinder \cite{livieri2007very}\end{tabular}                            & \begin{tabular}[c]{@{}l@{}}Widely used\\ clone detection\\ framework\end{tabular}   & ✓                             & ✓            & ✗            & ✗            \\ \hline
CCFinderSW \cite{semura2017ccfindersw}                                                                                                                 & \begin{tabular}[c]{@{}l@{}}ANTRL based\\ clone detection\end{tabular}               & ✓                             & ✗            & ✗            & ✓            \\ \hline
\begin{tabular}[c]{@{}l@{}}SourcererCC \cite{sajnani2016sourcerercc}\\ Oreo \cite{saini2018oreo}\end{tabular}                                         & \begin{tabular}[c]{@{}l@{}}These tools detect\\ all types of clones\end{tabular}    & ✓                             & ✗            & ✗            & ✓            \\ \hline
BigCloneEval \cite{svajlenko2016bigcloneeval}                                                                                                          & \begin{tabular}[c]{@{}l@{}}Large-scale clone\\ detection\end{tabular}               & ✓                             & ✓            & ✗            & ✗            \\ \hline
Deckard \cite{jiang2007deckard}                                                                                                                        & \begin{tabular}[c]{@{}l@{}}ANTRL Parse Tree\\ based Clone Detector\end{tabular}     & ✓                             & ✗            & ✓            & ✗            \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. For our tool, we decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio \cite{higo2013revisiting, kamalpriya2017enhancing}. However, instead of building a dependency graph, we build a similarity graph of statements linking to similar statements. We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is compatible with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). This is displayed in figure \ref{fig:clonerefactor}.

@@

On basis of this graph we detect clone classes. We challenge completed clone classes against the thresholds that are being used, and remove the clone classes that do not adhere to the thresholds.

@@

\section{Redefining type 1, 2 and type 3 clones}\label{chap:clonetypes}
In a 2007 survey by Roy et al \cite{roy2007survey} he defined four types of clones:

@@

\subsection{Type 1 clones} \label{chap:type1clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, does not yet indicate that they are actually identical. Even when textually equal, method calls can refer to different methods, type declarations can refer to different types and even variables can be of a different type. This could invalidate a refactoring opportunity for such a code fragment. Because of that, we recommend redefining type-1 clones for refactoring. We propose to:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} This way we can validate whether two method references, like method calls, are actually equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also of the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
\end{itemize}
